%\begin{savequote}[8cm]
%  ``Confession: I played golf. Confession: I hired a cleaning lady. Confession: I am owner of the Turbo nose-hair trimmer with optional ear-hair accessory.''
%  \qauthor{Dan Zevin}
%\end{savequote}
%\makeatletter

\chapter{Introduction}
\label{Introduction}

The \emph{fast Fourier transform (FFT)} has become one of the most important 
and widely used algorithms today. In 1965, Cooley and Tukey \cite{cotu} published
and publicized the first description of a fast algorithm, known nowadays as Cooley-Tukey 
FFT, for computing certain trigonometric sums, generally referred to as the 
\emph{discrete} or \emph{finite Fourier transform (DFT)}. In fact, developments since then, now
let us speak of a whole class of algorithms. The original Cooley-Tukey FFT
used a recursive scheme to split one large transform successively into
smaller transforms achieving an asymptotical complexity of 
$\bigo{N \log N}$ \emph{floating point operations (flops)} instead of
$\bigo{N^2}$ for a direct computation, where $N$ is the transform length. 
Efficient and highly optimized implementations also for the
multidimensional case are available \cite{fftw}.
Surprisingly, the idea of Cooley and Tukey was already described by Gauss around
1805 \cite{hejobu} interpolating the trajectories of the asteroids Pallas 
and Juno. But his work was not widely recognized and only published posthumously.
The wide areas of applications nowadays using FFT-techniques, including, for example, 
time-frequency analysis, signal-processing and the numerical solution of 
partial differential equations, underline the great importance and impact of fast
DFT algorithms.

Recently, algorithms have been developed for a more general type of discrete 
transforms, namely the \emph{nonuniform discrete Fourier transform (NDFT)}. 
While the DFT is a bijective and easily invertible linear mapping of $N$ 
ingoing coefficients to $N$ outgoing coefficients, corresponding to
uniformly distributed nodes on an interval, or in higher dimensions on 
the multidimensional Torus, the NDFT generalizes
discrete Fourier sums for arbitrary node distributions, hence the name
'nonuniform'. Fast algorithms have been described in several papers 
(see \cite{postta01} and the references therein) and a C subroutine library is available 
(\cite{kupo02C}).

A different line of generalization leaves the 'classical' setting of 
Fourier series on a multidimensional torus and analogously describes
Fourier series on different geometries like the surface of a
multidimensional sphere. Particularly, the unit sphere $\twosphere$ 
embedded into $\R^3$ has practical relevance in quite a wide range of 
applications, in many cases due to its correspondence to the surface 
of the earth. Fields of interest are numerical analysis in
geo-sciences, for example in the computation of climate models for
weather forecasts or regarding the gravitational potential of the 
earth. Moreover, data distributed on the surface of a sphere arises
in many different applications in a natural way. Examples are
molecular dynamics for protein docking problems and 
gamma-cameras for applications in computed tomography.
Unfortunately, the spherical setting differs from the 'classical' 
one in that the numerical treatment of many computational problems
is quite more challenging. Although there exists a rich theory for 
analysis on the sphere, fast algorithms allowing for efficient and
reliable discrete transforms in terms of so-called 
\emph{spherical surface functions}, the spherical counterpart of the 
Fourier basis of complex exponentials on the torus,
had not been developed until a first paper by Driscoll and Healy 
\cite{drhe}.
 
During the past years, major progress has been made and several 
different techniques have been proposed (see \cite{HeRoKoMo}, 
\cite{suta}, \cite{roty}). The basic idea for the 
methods presented in this text is to convert the discrete spherical 
Fourier transform into a 'classical' two-dimensional discrete 
Fourier transform. A careful analysis
of numerical instabilities involved has proven to be of major 
importance.

The aim of this text is to describe particular algorithms for fast discrete
spherical Fourier transforms. Chapter \ref{Basics} gives a brief 
introduction to Fourier analysis on the sphere $\twosphere$ and 
provides fundamental facts and results. We refer to the literature for 
details.

In Chapter \ref{DSFT}, algorithms for the computation of discrete
spherical Fourier transforms for arbitrary nodes on the sphere 
are presented. We first describe the so-called 'direct' algorithms 
and mention how spherical Fourier coefficients might be computed 
by certain quadrature formulae: Starting with a simple algorithm 
for the evaluation of a \emph{spherical Fourier sum} at arbitrary 
nodes, we derive the matrix factorization corresponding to the 
algorithm translated into matrix-vector notation. Based on the 
easily obtained \emph{adjoint matrix factorization}, we 
describe the so-called \emph{adjoint algorithm}. In conjunction
with appropriate quadrature formulae, we are able to reconstruct
spherical Fourier coefficients from function samples on the sphere.
We then present corresponding 'fast' algorithms which employ a
\emph{discrete polynomial transform}. The work is based on 
previous papers concerning an algorithm for 
the fast approximative evaluation of \emph{spherical Fourier sums} 
in terms of spherical surface functions at arbitrary nodes on the 
sphere. The main contribution of this text is the derivation of the corresponding
matrix factorization along with the construction of the 
\emph{adjoint algorithm}. All algorithms were implemented and tests 
of the 'fast' algorithms with respect to numerical stability and 
runtime performance are provided.
Chapter \ref{Applications} finally presents two applications of the 
proposed algorithms: 

We propose a new method for the fast 
approximative summation of \emph{spherical radial basis functions}
and derive theoretical error bounds for some example functions. The 
theoretical results are verified by numerical tests and the runtime
performance of the presented algorithms is compared with a direct
and exact implementation.

We finally treat the generally ill-posed problem of reconstructing 
spherical Fourier coefficients from scattered data. We formulate
the problem as a \emph{weighted linear least squares problem}
and a \emph{weighted optimal interpolation problem} using
\emph{normal equations} and describe how the fast algorithms
for discrete spherical Fourier transforms can be used to obtain 
a solution by means of standard iterative algorithms for systems 
of linear equations. Examples are given for the former case using 
data of the earth's gravitational potential.

